Alex's Notes and Text Fragments for the JSS Manuscript

### 2021-05-25 paper/paper.Rmd
using caligraphic letters for sets

In cross-validation for predictive model assessment, we consider the following formal setting.
We are interested in predicting a numerical or categorical response $y$ of an object or instance using a feature vector $\mathbf{x} = (x^{(1)}, \ldots, x^{(p)})\in\mathbb{R}^p$ and a model $\hat{f}_L$ that has been trained on a learning sample $L = \{(y_i, \mathbf{x}_i),\ i = 1, ..., n\}$. Our goal is to estimate the expected value of the performance of $\hat{f}_L$,
\[
\mathit{perf(\hat{f}_L)} := E(l(Y,\hat{f}_L(X))),
\]
where $l$ is a real-valued loss function, and the expected value is with respect the probability distribution of $X$, the features of an instance $(Y,X)$ drawn randomly from the underlying population.
This is referred to as the *actual* or *conditional* performance measure as it is conditional on $L$ [@milligan2000].
The loss function can take a variety of forms such as the misclassification error $I(Y\neq\hat{f}_L(X))$ in classification, or the squared error $(Y-\hat{f}_L(X))^2$ in regression, among many other possible measures.
The choice of the performance measure is equally critical as the choice of the estimation procedure, but it is beyond the scope of this contribution to discuss performance metrics for regression and classification (see e.g., @hand1997 for classification, and ref?? for regression tasks).

Nevertheless, since we only have a sample $T$ of test data drawn from the population, we can realistically only *estimate* the conditional performance of $\hat{f}_L$:
\[
\widehat{\mathit{perf}}_T(\hat{f}_L) = \frac{1}{|T|}\sum_{i=1}^{|T|}l(Y,\hat{f}_L(X)).
\]
This representation as a point estimator for $\mathit{perf(\hat{f}_L)}$ underlines the importance of using a random sample for performance estimation to avoid estimation bias. Other estimators than the simple mean may be required when $T$ is not a simple random sample, for instance a stratified random sample [e.g., @thompson2012].
As always, judgment sampling may lead to uncontrollable bias.

Since the re-use of the learning sample $L$ for testing, $T=L$, would result in the over-optimistic *resubstitution* or *apparent* model performance, cross-validation partitions the sample $L$ into disjoint training and test sets.
Specifically, $L$ is split into $k$ partitions,
\[
L = L_1 \cup \ldots \cup L_k,\qquad L_i\cap L_j = \emptyset\quad \textrm{for all}\ i\neq j,
\]
and a model $\hat{f}_{(i)}$ is fitted on $L_{(i)} := L\setminus L_i$, while $L_i$ is withheld for testing.
This is repeated for $i=1,\ldots,k$ in order to effectively use the entire sample for testing, while keeping training and test sets disjoint at all times.
The $k$-fold CV estimator can therefore be written as
\[
\widehat{\mathit{perf}}_{L, CV}(f) := \frac{1}{k}\sum_{i=1}^k\widehat{\mathit{perf}}_{L_i}(\hat{f}_{L_{(i)}}),
\]
where $f$ is a machine-learning algorithm, i.e. a mapping that trains a model $\hat{f}_S$ using any suitable training sample $S$.
The use of $k=5$ or $k=10$ folds is most commonly seen in practice, and these preferences are also supported by theory [ref??].
The $k$-fold CV estimator of model performance is a nearly unbiased estimator of the conditional performance measure when the observations were drawn independently [@efron1983].
Since $\widehat{\mathit{perf}}_{L, CV}(f)$ still depends on the particular partitioning chosen for $L$, it is sometimes recommended to repeat the estimation using different random partitionings ($r$-repeated $k$-fold cross-validation, e.g. $k=100$; ref??).

>>>>

In cross-validation for predictive model assessment, we consider the following formal setting.
We are interested in predicting a numerical or categorical response $y$ of an object or instance using a feature vector $\mathbf{x} = (x^{(1)}, \ldots, x^{(p)})\in\mathbb{R}^p$ and a model $\hat{f}_\mathcal{L}$ that has been trained on a learning sample $\mathcal{L} = \{(y_i, \mathbf{x}_i),\ i = 1, ..., n\}$. Our goal is to estimate the expected value of the performance of $\hat{f}_\mathcal{L}$,
\[
\mathit{perf(\hat{f}_\mathcal{L})} := E(l(Y,\hat{f}_\mathcal{L}(X))),
\]
where $l$ is a real-valued loss function, and the expected value is with respect the probability distribution of $X$, the features of an instance $(Y,X)$ drawn randomly from the underlying population.
This is referred to as the *actual* or *conditional* performance measure as it is conditional on $L$ [@milligan2000].
The loss function can take a variety of forms such as the misclassification error $I(Y\neq\hat{f}_\mathcal{L}(X))$ in classification, or the squared error $(Y-\hat{f}_\mathcal{L}(X))^2$ in regression, among many other possible measures.
The choice of the performance measure is equally critical as the choice of the estimation procedure, but it is beyond the scope of this contribution to discuss performance metrics for regression and classification (see e.g., @hand1997 for classification, and ref?? for regression tasks).

Nevertheless, since we only have a sample $\mathcal{T}$ of test data drawn from the population, we can realistically only *estimate* the conditional performance of $\hat{f}_\mathcal{L}$:
\[
\widehat{\mathit{perf}}_T(\hat{f}_\mathcal{L}) = \frac{1}{|\mathcal{T}|}\sum_{i=1}^{|\mathcal{T}|}l(Y,\hat{f}_\mathcal{L}(X)).
\]
This representation as a point estimator for $\mathit{perf(\hat{f}_\mathcal{L})}$ underlines the importance of using a random sample for performance estimation to avoid estimation bias. Other estimators than the simple mean may be required when $\mathcal{T}$ is not a simple random sample, for instance a stratified random sample [e.g., @thompson2012].
As always, judgment sampling may lead to uncontrollable bias.

Since the re-use of the learning sample $\mathcal{L}$ for testing, $\mathcal{T}=\mathcal{L}$, would result in the over-optimistic *resubstitution* or *apparent* model performance, cross-validation partitions the sample $\mathcal{L}$ into disjoint training and test sets.
Specifically, $\mathcal{L}$ is split into $k$ partitions,
\[
\mathcal{L} = \mathcal{L}_1 \cup \ldots \cup \mathcal{L}_k,\qquad \mathcal{L}_i\cap \mathcal{L}_j = \emptyset\quad \textrm{for all}\ i\neq j,
\]
and a model $\hat{f}_{(i)}$ is fitted on $\mathcal{L}_{(i)} := \mathcal{L}\setminus \mathcal{L}_i$, while $\mathcal{L}_i$ is withheld for testing.
This is repeated for $i=1,\ldots,k$ in order to effectively use the entire sample for testing, while keeping training and test sets disjoint at all times.
The $k$-fold CV estimator can therefore be written as
\[
\widehat{\mathit{perf}}_{\mathcal{L}, CV}(f) := \frac{1}{k}\sum_{i=1}^k\widehat{\mathit{perf}}_{\mathcal{L}_i}(\hat{f}_{\mathcal{L}_{(i)}}),
\]
where $f$ is a machine-learning algorithm, i.e. a mapping that trains a model $\hat{f}_\mathcal{S}$ using any suitable training sample $\mathcal{S}$.
The use of $k=5$ or $k=10$ folds is most commonly seen in practice, and these preferences are also supported by theory [ref??].
The $k$-fold CV estimator of model performance is a nearly unbiased estimator of the conditional performance measure when the observations were drawn independently [@efron1983].
Since $\widehat{\mathit{perf}}_{\mathcal{L}, CV}(f)$ still depends on the particular partitioning chosen for $L$, it is sometimes recommended to repeat the estimation using different random partitionings ($r$-repeated $k$-fold cross-validation, e.g. $k=100$; ref??).
